---
title: "Lab9"
author: "James McKinlay"
date: "2022-11-29"
output: html_document
---

```{r}
library(mlbench)
data("Ionosphere")
library(cluster)
library(tidyverse)
library(ggplot2)
```

```{r}
head(Ionosphere)
```

1)

```{r}
eucl <- dist(Ionosphere[,c(-1,-2,-35)], method="euclidean")
manh <- dist(Ionosphere[,c(-1,-2,-35)], method="manhattan")
mink <- dist(Ionosphere[,c(-1,-2,-35)], method="minkowski")
gower <- daisy(Ionosphere[,-35], metric="gower")
```

```{r}
summary(eucl)
```
```{r}
summary(manh)
```

```{r}
summary(mink)
```

```{r}
summary(gower)
```

2)
```{r}
eucl2 <- dist(Ionosphere[,-1:-2], method="euclidean")
manh2 <- dist(Ionosphere[,-1:-2], method="manhattan")
mink2 <- dist(Ionosphere[,-1:-2], method="minkowski")
gower2 <- daisy(Ionosphere, metric="gower")
```

```{r}
summary(eucl2)
```
 Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  0.000   2.864   4.212   4.050   5.247   9.889 



```{r}
summary(manh2)
```
  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   0.00   13.23   19.75   19.18   25.01   50.44 
   

```{r}
summary(mink2)
```
 Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  0.000   2.864   4.212   4.050   5.247   9.889 


```{r}
summary(gower2)
```

61425 dissimilarities, summarized :
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
 0.0000  0.1896  0.2842  0.2769  0.3612  0.7353 
Metric :  mixed ;  Types = N, N, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I, I 
Number of objects : 351


All metrics are very similar with regards to 1st quartile, median, mean, 3rd quartile, and max (all mins are the same at 0)


3)

```{r}
eucl <- as.matrix(eucl)
eucl2 <- as.matrix(eucl2)

manh <- as.matrix(manh)
manh2 <- as.matrix(manh2)

mink <- as.matrix(mink)
mink2 <- as.matrix(mink2)

gower <- as.matrix(gower)
gower2 <- as.matrix(gower2)
```

```{r}
eucl_first_row <- eucl[1,]
eucl_first_row <- sort(eucl_first_row)

head(eucl_first_row)
tail(eucl_first_row)
```

For the euclidean matrix, the most similar cases in the first row are cases: 33, 182, 3, 9, and 145.  The most dissimilar cases are cases: 22, 163, 207, 54, 189, and 18.

```{r}
manh_first_row <- manh[1,]
manh_first_row <- sort(manh_first_row)

head(manh_first_row)
tail(manh_first_row)
```
For the manhattan matrix, the most similar cases are cases: 182, 33, 3, 145, and 202.  The most dissimilar cases are cases: 163, 221, 22, 189, 54, 18
```{r}
mink_first_row <- mink[1,]
mink_first_row <- sort(mink_first_row)

head(mink_first_row)
tail(mink_first_row)
```
For the Minkowski matrix, the most similar cases are cases: 33, 182, 3, 9, and 145, while the most dissimilar are cases: 22, 163, 207, 54, 189, and 18
```{r}
gower_first_row <- gower[1,]
gower_first_row <- sort(gower_first_row)

head(gower_first_row)
tail(gower_first_row)
```
For the Gower matrix, the most similar cases are cases: 182, 33, 3, 145, and 202, while the most dissimilar cases are cases: 20, 221, 54, 22, 189, and 18.


4)

```{r}
eucl <- dist(Ionosphere[,c(-1,-2,-35)], method="euclidean")
iono_without_class <- Ionosphere[, c(-1,-2,-35)]

set.seed(123)
kmeans_iono1 <- kmeans(iono_without_class, centers=15)
kmeans_iono2 <- kmeans(iono_without_class, centers=2)
kmeans_iono3 <- kmeans(iono_without_class, centers=3)
kmeans_iono10 <- kmeans(iono_without_class, centers=10)
kmeans_iono20 <- kmeans(iono_without_class, centers=20)


sil <- silhouette(kmeans_iono1$cluster, dist=eucl)
sil2 <- silhouette(kmeans_iono2$cluster, dist=eucl)
sil3 <- silhouette(kmeans_iono3$cluster, dist=eucl)
sil4 <- silhouette(kmeans_iono10$cluster, dist=eucl)
sil5 <- silhouette(kmeans_iono20$cluster, dist=eucl)

plot(sil, border=NA)
plot(sil2, border=NA)
plot(sil3, border=NA)
plot(sil4, border=NA)
plot(sil5, border=NA)

mean1<-mean(sil[,3])

mean2<-mean(sil2[,3])

mean3<-mean(sil3[,3])

mean4<-mean(sil4[,3])

mean5<-mean(sil5[,3])
```

5)
```{r}
avg <- c()
for(k in 2:6){
  clust <- kmeans(iono_without_class, centers=k, iter.max=100)
  silo <- silhouette(clust$cluster, eucl)
  avg <- c(avg, mean(silo[,3]))
}
```

```{r}
ggplot(data.frame(nClus=2:6,Silh=avg), aes(x=nClus,y=Silh)) + geom_point(size=3,color="red") + geom_line() +
xlab("Num                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                       .Clusters") + ylab("Silo.Coef.")
```

